An iterated Grothendieck construction of semi-cubical types

### 2022-02-23 paper/paper.tex
art.cls: switch to brand new art class

\documentclass[10pt]{amsart}

\usepackage[allbordercolors={.192157 .76862 .28627}]{hyperref} % Emerald Green
\usepackage{amsthm, amssymb, amsmath, mathtools, amsaddr, fancyhdr, bookmark, fontspec, xargs, multirow, polyglossia, luacode, minted, color, soulutf8, tcolorbox, tabularx, tikz-cd, booktabs, setspace, makecell, caption, mathrsfs, xpatch, lmodern}
\usepackage[hmargin=0.4in, vmargin=0.8in, headheight=10pt]{geometry}

@@

% Objects and morphisms
\newcommand{\obj}{\ensuremath{\mathsf{obj}}}
\newcommand{\mor}{\ensuremath{\mathsf{mor}}}

% Simplicial sets
\newcommand{\Set}{\ensuremath{\mathsf{Set}}}
\newcommand{\op}{\ensuremath{\mathsf{op}}}
\newcommand{\Simplex}[1]{\ensuremath{\boldsymbol{\Delta^{#1}}}}
\newcommand{\DeltaHat}{\ensuremath{\hat{\boldsymbol{\Delta}}}}
\newcommand{\SSet}{\ensuremath{\mathsf{Set}_{\boldsymbol{\Delta}}}}
\newcommand{\sq}[1]{\ensuremath{\mathsf{[#1]}}}

% Cubical sets
\newcommand{\Cube}[1]{\ensuremath{\boldsymbol{\square^{#1}}}}
\newcommand{\I}[1]{\ensuremath{\mathsf{I}^{#1}}}
\newcommand{\CSet}{\ensuremath{\mathsf{Set}_{\boldsymbol{\square}}}}

@@

% Some abbreviations
\newcommand{\D}{\ensuremath{(D)}}
\newcommand{\hdD}{\ensuremath{(D.1)}}
\newcommand{\tlD}{\ensuremath{(D.2)}}
\renewcommand{\d}{\ensuremath{(d)}}
\newcommand{\E}{\ensuremath{(E)}}
\renewcommand{\l}{\ensuremath{(l)}}
\renewcommand{\c}{\ensuremath{(c)}}

@@

% Use lettered footnotes
\renewcommand{\thefootnote}{\alph{footnote}}

% amsthm theorems
\swapnumbers
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{terminology}[definition]{Terminology}
\newtheorem{notation}[definition]{Notation}
\newtheorem{proposition}[definition]{Proposition}
\newtheorem{lemma}[definition]{Lemma}
\newtheorem{theorem}[definition]{Theorem}
\newtheorem{corollary}[definition]{Corollary}
\newtheorem{remark}[definition]{Remark}
\newtheorem{example}[definition]{Example}
\newtheorem{caution}[definition]{Caution}
\newtheorem{discussion}[definition]{Discussion}

\usepackage{thmtools}

% Fancy headers
\pagestyle{fancy} % choose the "fancy" pagestyle
\fancyhf{}        % clear all headers and footers
% Now set the headers
\fancyhead[LE,RO]{\footnotesize\thepage}
\fancyhead[LO,RE]{\footnotesize\leftmark}
\makeatletter
\xapptocmd{\@sect}{\csname #1mark\endcsname{#7}}{}{}
\makeatother

% Subsection centering
\makeatletter
\def\subsection{\@startsection{subsection}{2}
  \z@{.5\linespacing\@plus.7\linespacing}{.6\baselineskip}{\centering}}
\makeatother

% Nested numbering of sections and subsections
\makeatletter
\newif\ifsection
\newif\ifsubsection
\preto\section{\sectiontrue\subsectionfalse}
\preto\subsection{\sectionfalse\subsectiontrue}
\patchcmd{\@xsect}% <cmd>
  {\ignorespaces}% <search>
  {\ifsubsection
   \numberwithin{definition}{subsection}
   \else
   \numberwithin{definition}{section}
   \fi
   \setcounter{definition}{0}\relax\ignorespaces}% <replace>
  {}{}% <success><failure>
\makeatother

% Fancy headers
\pagestyle{fancy} % choose the "fancy" pagestyle
\fancyhf{}        % clear all headers and footers
% Now set the headers
\fancyhead[LE,RO]{\footnotesize\thepage}
\fancyhead[LO,RE]{\footnotesize\leftmark}
\makeatletter
\xapptocmd{\@sect}{\csname #1mark\endcsname{#7}}{}{}
\makeatother

% Subsection centering
\makeatletter
\def\subsection{\@startsection{subsection}{2}
  \z@{.5\linespacing\@plus.7\linespacing}{.6\baselineskip}{\centering}}
\makeatother

@@

% Fonts
\setmainfont{Avenir Next}
\setmonofont{Source Code Pro}
\newfontfamily\mainfont{Avenir Next}
\defaultfontfeatures{Ligatures=TeX, Scale=MatchUppercase}

\title{An iterated Grothendieck construction of semi-cubical types}
\author{Hugo Herbelin}
\address{Institut de Recherche en Informatique Fondamentale}
\author{Ramkumar Ramachandra}
\address{Université de Paris}

\begin{document}
\begin{abstract}
  A proposed model of \emph{homotopy type theory} built via a \emph{reverse iterated Grothendieck construction} process formalizing semi-cubical sets in dependent type theory with \emph{extensional equality}. The accompanying artifact is a Coq formalization that uses highly sophisticated rewriting techniques and pushes the boundary of proof assistant technology.
\end{abstract}
\maketitle
\tableofcontents
\newpage

@@

  \begin{equation*}
    \SSet := \Set^{\Simplex{}^\op}
  \end{equation*}

@@

  \begin{align*}
    \obj(\SSet)            & := \text{functors} \; \Simplex{}^\op \rightarrow \Set      \\
    \mor(\SSet)            & := \text{natural transforms between the functors}          \\
    \text{composition law} & := \text{the usual composition of natural transformations} \\
  \end{align*}
\end{definition}

@@

\subsection{Motivation for \texorpdfstring{\DeltaHat}{delta sets}}
As hinted earlier, simplicial sets are ill-suited for machine-based representation, due to the infinite number of degenerate simplices that are uninteresting from the point of view of cubical type theory. To remedy this problem, let us instead use \DeltaHat, and proceed towards building, what is called \emph{semi-cubical sets} in literature. $\DeltaHat$ is identical to \Simplex{}, but for the fact that the maps $\sq{m} \rightarrow \sq{n}$ are \emph{strictly} order-preserving. The degeneracies then vanish, and we're left with the task of defining gluing conditions solely on the basis of face maps. In mathematical literature, there is little interest in studying \DeltaHat, since delta maps are not well-defined, and constructing a $\Set^{\DeltaHat^{op}}$ is inelegant.

>>>>

\documentclass[10pt]{art}

\usepackage{multirow, minted, tabularx, booktabs, setspace, makecell, caption}
\usepackage[hmargin=0.4in, vmargin=0.8in, headheight=10pt]{geometry}

% Simplicial and Cubical sets
\newcommand{\DeltaHat}{\ensuremath{\hat{\boldsymbol{\Delta}}}}
\newcommand{\Cube}[1]{\ensuremath{\boldsymbol{\square^{#1}}}}
\renewcommand{\I}[1]{\ensuremath{\mathsf{I}^{#1}}}
\newcommand{\CSet}{\ensuremath{\mathsf{Set}_{\boldsymbol{\square}}}}

@@

@@

% Some abbreviations
\renewcommand{\D}{\ensuremath{(D)}}
\newcommand{\hdD}{\ensuremath{(D.1)}}
\newcommand{\tlD}{\ensuremath{(D.2)}}
\renewcommand{\d}{\ensuremath{(d)}}
\newcommand{\E}{\ensuremath{(E)}}
\renewcommand{\l}{\ensuremath{(l)}}
\renewcommand{\c}{\ensuremath{(c)}}

% Fancy headers
\pagestyle{fancy}
\fancyhf{}
\fancyhead[R]{\footnotesize\textcolor{gray80}{{\thepage / \pageref{LastPage}}}}
\renewcommand{\sectionmark}[1]{\markboth{}{\thesection.\ #1}}
\fancyhead[L]{\footnotesize\textcolor{gray80}{{\MakeUppercase{\rightmark}}}}

% Section and subsection styles
\renewcommand{\thesection}{\Roman{section}}
\titleformat{\section}{\centering\scshape\Large\color{raspeberry}}{\thesection}{0.7em}{}
\titleformat{\subsection}{\strongenv\large\color{gray80}}{\thesubsection}{0.7em}{}

% Pack enumerate items and bibitems
\setlist{nolistsep}
\setlength{\bibitemsep}{.2\baselineskip plus .05\baselineskip minus .05\baselineskip}

@@

@@

% Fonts
\setmainfont{Avenir Next}
\setmonofont{Source Code Pro}
\defaultfontfeatures{Ligatures=TeX, Scale=MatchUppercase}

\title{An iterated Grothendieck construction of semi-cubical types}
\author{
  \textcolor{gray80}{Hugo Herbelin} \\
  \itshape \textcolor{gray80}{IRIF}
  \and
  \textcolor{gray80}{Ramkumar Ramachandra} \\
  \itshape \textcolor{gray80}{Université de Paris}
}
\date{}

\begin{document}
\thispagestyle{empty}
\maketitle
\begin{abstract}
  A proposed model of \emph{homotopy type theory} built via a \emph{reverse iterated Grothendieck construction} process formalizing semi-cubical sets in dependent type theory with \emph{extensional equality}. The accompanying artifact is a Coq formalization that uses highly sophisticated rewriting techniques and pushes the boundary of proof assistant technology.
\end{abstract}
\tableofcontents
\newpage

@@

  \begin{equation*}
    \SSet := \Set^{\op{\Simplex{}}}
  \end{equation*}

@@

  \begin{align*}
    \obj(\SSet)            & := \text{functors} \; \op{\Simplex{}} \rightarrow \Set     \\
    \mor(\SSet)            & := \text{natural transforms between the functors}          \\
    \text{composition law} & := \text{the usual composition of natural transformations} \\
  \end{align*}
\end{definition}

@@

\subsection{Motivation for \texorpdfstring{\DeltaHat}{delta sets}}
As hinted earlier, simplicial sets are ill-suited for machine-based representation, due to the infinite number of degenerate simplices that are uninteresting from the point of view of cubical type theory. To remedy this problem, let us instead use \DeltaHat, and proceed towards building, what is called \emph{semi-cubical sets} in literature. $\DeltaHat$ is identical to \Simplex{}, but for the fact that the maps $\sq{m} \rightarrow \sq{n}$ are \emph{strictly} order-preserving. The degeneracies then vanish, and we're left with the task of defining gluing conditions solely on the basis of face maps. In mathematical literature, there is little interest in studying \DeltaHat, since delta maps are not well-defined, and constructing a $\Set^{\op{\DeltaHat}}$ is inelegant.
